---
title: "460 Model_RandomForest"
author: "Geyi Liu"
date: "2021/1/10"
output:
  html_document: default
  pdf_document: default
---

Introduction:
https://corporatefinanceinstitute.com/resources/knowledge/other/random-forest/
https://pubs.acs.org/doi/10.1021/ci034160g


The essence of random forest algorithm is classifier ensemble algorithm based on decision tree. Compared with neural network, random forest reduces the amount of computation and improves the prediction accuracy. Moreover, the algorithm is insensitive to collinearity and robust to missing data and unbalanced data, which can be well adapted to thousands of explanatory variable data sets.


```{r}
library(pROC) # Draw ROC curve
library(randomForest)
library(stringr)
library(dplyr)
data <- read.csv("Data.csv", stringsAsFactors=FALSE)
data$LAB <- as.factor(data$LAB) # convert LAB to be a factor
data[is.na(data)] <- 0 # replace NAs with zero
data_model <- data
data_model$X <- NULL # drop identifier column
data_model$LAB <- NULL # drop non-numeric LAB column
data_model <- data_model[1:nrow(data_model),477:ncol(data_model)]
```

```{r}
a<-ncol(data_model) # number of columns
b<-nrow(data_model)*0.8 # 80% number of rows
c<-c()
# Delete columns with more than 80% zeros
for(i in 1:a){
  # print(sum(data_model[,i]==0)) # Number of 0 per column
  if( sum(data_model[,i]==0)>=b ){
    c=append(c,i)
  }
  }
print(c) # Columns to be deleted
Data_remove <- data_model[,-c]
Data_remove$Pass <- as.numeric(Data_remove$Response >= 50)
Data_remove$Pass <- as.factor(Data_remove$Pass)
# Replace Response with Pass
Data_remove <- subset(Data_remove, select = -c(Response))
```


```{r}
set.seed(100)
# Divide the data set  into training set and test set, the ratio is 7:3
train_sub <- sample(nrow(Data_remove),7/10*nrow(Data_remove))
train_data <- Data_remove[train_sub,]
test_data <- Data_remove[-train_sub,]
```


# Some important parameters of randomForest()

ntree: the number of trees in the forest. The default is 500
mtry: number of features used per tree
Importance: whether to calculate theimportance of variables. The default value is false
Proximity: whether to calculate the similarity between observations

```{r}
# There are many variables, so the operation time will be long
set.seed(100)
rf <- randomForest(Pass ~ ., data=train_data,importance=TRUE, proximity=TRUE,type=classification)
print(rf)
```



```{r}
print(rf)
TP<-rf$confusion[4]
TN<-rf$confusion[1]
FN<-rf$confusion[2]
FP<-rf$confusion[3]
print(FP)
# Recall
Rec <-TP/(TP+FN)
# Precision
Pre<-TP/(TP+FP)
# Accuracy
Acc<-(TP+TN)/(TP+FN+TN+FP)
# F1 score
F1 <- 2*Acc*Rec/(Acc+Rec)
```

View the importance of variables:
Mean decrease accuracy: the decrease of accuracy after replacing this variable.
Mean decrease Gini: the decrease of Gini coefficient after variable replacement. The larger the value, the more important the variable is.


```{r}
# importance()
imp <- as.data.frame(rf$importance)
# Check which variables are important
best <- subset(imp,imp$MeanDecreaseGini>10)
```
Group1_11,Group1_9,Group1_2 are the most important factors that affect whether the product passes the test.

Group1 and Group2  are the three most influential groups.


```{r}
# Output the top 15 important variables
varImpPlot(rf, main = "variable importance",n.var=15)
```

```{r}
# Predicting the test sets
pre_ran <- predict(rf,newdata=test_data)
# Integrating real and predicted values
obs_p_ran <- data.frame(prob=pre_ran,obs=test_data$Pass)
# Output confusion matrix
table(test_data$Pass,pre_ran,dnn=c("Real","Predict"))
# Draw ROC curve
ran_roc <- roc(test_data$Pass,as.numeric(pre_ran))

plot(ran_roc, print.auc=TRUE, auc.polygon=TRUE, grid=c(0.1, 0.2),grid.col=c("green", "red"), max.auc.polygon=TRUE,auc.polygon.col="skyblue", print.thres=TRUE,main='ROC curve of random forest model')
```


```{r}
pred_out_1<-predict(object=rf,newdata=test_data,type="prob")
table <- table(test_data$Pass,pre_ran)
print(table)
# diag(table): Extract the values on the diagonal of the matrix
sum(diag(table))/sum(table)  # Prediction accuracy
```

AUC（Area Under Curve: ) is used to indicate the accuracy of prediction. The higher the AUC (the larger the area under the curve), the higher the accuracy of prediction.

mtry and ntree will affect the results of the model, but the influence is not very significant. 

The most important variables obtained by the random forest model are inconsistent with those obtained by the linear regression model. Classification model and linear regression model have different results because of their different principles. 

The results of random forest can be compared with the results of other classification models.

